Late Pleistocene stone artefact technology at Mau A, Yen Bai, Norther Vietnam

### 2022-02-25 analysis/paper/paper.Rmd
Reordered paragraphs in the methods section
Revised and added new citations for univariate k-means

@@

Dorsal cortex percentages discretized upon standard cut points were used to create reduction categories. Flakes with no dorsal cortex were categorized as tertiary flakes, those with up to a dorsal cortex percentage of 50% were considered secondary and those with a dorsal cortex value greater than 50% were considered primary (Bradbury & Carr 1995). Mass clusters were generated using univariate k-means, with the end result being 6 mass clusters, with mass cluster 1 having the least mass on average and 6 having the greatest mass on average. Univariate k-means is a non-hierarchical clustering technique that creates a set number of groups equal to k, with points being assigned to groups in such a manner that the total distances between each point and mean of the group in minimized. For more information on univariate k-means, Papageorgiou (2020) examines the various different statistical techniques (including k-means) that can be applied in an archaeological setting, specifically in the study of ceramics.

@@

Papageorgiou, I. (2020). Ceramic investigation: How to perform statistical analyses. Archaeological and Anthropological Sciences, 12(9), 210. https://doi.org/10.1007/s12520-020-01142-x

@@

>>>>

  Dorsal cortex percentages discretized upon standard cut points were used to create reduction categories. Flakes with no dorsal cortex were categorized as tertiary flakes, those with up to a dorsal cortex percentage of 50% were considered secondary and those with a dorsal cortex value greater than 50% were considered primary (Bradbury & Carr 1995). Mass clusters were generated using univariate k-means, with the end result being 6 mass clusters, with mass cluster 1 having the least mass on average and 6 having the greatest mass on average.

  Univariate k-means is a non-hierarchical clustering technique that creates a set number of groups equal to k, with points being assigned to groups in such a manner that the total distances between each point and mean of the group in minimized. We used the Ckmeans.1d.dp package in R to perform our univariate k-means analysis (Wang & Song, 2011). The Ckmeans.1d.dp has been used for univariate clustering in applications outside of archaeology prior to our work (Song & Zhong, 2020).

@@

@@

Song, M., & Zhong, H. (2020). Efficient weighted univariate clustering maps outstanding dysregulated genomic zones in human cancers. Bioinformatics, 36(20), 5027–5036. https://doi.org/10.1093/bioinformatics/btaa613

@@

Wang, H., & Song, M. (2011). Ckmeans.1d.dp: Optimal k-means Clustering in One Dimension by Dynamic Programming. The R Journal, 3(2), 29–33.
